# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
unzip('activity.zip')
activity <- read.csv('activity.csv')
```      


## What is mean total number of steps taken per day?

The distribution of steps taken for sampled days is shown in the histogram below.
```{r}
daily.agg <- aggregate(steps ~ date, activity, FUN=sum)
hist(daily.agg$steps, xlab="Steps per day")
```


```{r}
options(scipen=99)
print(steps.mean <- mean(daily.agg$steps))
print(steps.median <- median(daily.agg$steps))
```

The mean number of steps taken was `r steps.mean`, and median was `r steps.median`.

## What is the average daily activity pattern?

We can also look at the average steps taken for each interval as a time series.
```{r}
interval.agg <- aggregate(steps ~ interval, activity, FUN=mean)
plot( steps ~ interval, interval.agg, type='l', main="Steps taken over intervals throughout the day")
```

```{r}
interval.max <- interval.agg[interval.agg$steps == max(interval.agg$steps),]
interval.max
```
The maximum average steps `r interval.max$steps` occured at interval `r interval.max$interval`.

## Imputing missing values

Number of missing values:
```{r}
sum(is.na(activity$steps))
```
We can then build a loess curve to estimate missing values using interval.
```{r}
model <- loess(steps ~ interval, activity)
imputed_values <- predict(model, activity[is.na(activity$steps),], type='response')
activity[is.na(activity$steps),]$steps <- imputed_values
```

New metrics after imputing missing values:
```{r}
daily.agg <- aggregate(steps ~ date, activity, FUN=sum)
hist(daily.agg$steps, xlab="Steps per day")
```

Mean steps per day after imputing missing values:
```{r}
mean(daily.agg$steps)
```

Median steps per day after imputing missing values:
```{r}
median(daily.agg$steps)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lattice)
library(plyr)
activity$Weekend <-  as.factor(weekdays(as.Date(activity$date)) %in% c("Saturday", "Sunday"))
activity$Weekend <- revalue(activity$Weekend, c("FALSE"="weekday", "TRUE"="weekend"))
interval.agg <- aggregate(steps ~ interval + Weekend, activity, FUN=mean)

xyplot(steps ~ interval|Weekend, interval.agg, layout=c(1,2), type='l')
```